Further Modeling of $q_{95}$ Windows for the Suppression of Edge Localized Modes by Resonant Magnetic Perturbations in the DIII-D Tokamak

### 2020-09-09 Papers/Paper2/Paper.tex
First draft of Paper2

\title {Modelling $q_{95}$ Windows for the Suppression of Edge Localized Modes by Resonant Magnetic Perturbations in the DIII-D Tokamak: Relaxation of the No-Slip
Constraint}

@@

%\begin{abstract}
%\end{abstract}

\section{Introduction}
In the context of tokamak plasmas, a resonant magnetic perturbation (RMP) refers to an externally
generated, (usually) static, helical magnetic perturbation that resonates at
one or more rational surfaces within the plasma. Resonant magnetic perturbations have been successfully
employed to compensate error fields (i.e., accidentally produced RMP), and, thereby,  to prevent the formation of
locked modes in a wide variety of different types of tokamak discharge. \cite{error1,error2,error3} RMPs have also been used to either control or
mitigate edge-localized-modes (ELMs) in H-mode tokamak discharges.\cite{elm1,elm2,elm3,elm4,elm5,elm6}

By far the most efficient manner in which to model the response of a tokamak plasma to an RMP is
to employ an ``asymptotic matching'' approach.\cite{fkr,rfa, am1,am2,am3} 
According to such an approach, the response of the plasma to the applied RMP is governed by a combination of flux-freezing and
perturbed force balance---this combination is usually referred to as ``marginally-stable ideal-magnetohydrodynamics (MHD)", which is a misnomer because MHD {\em per se}\/ plays no
role---everywhere in the plasma apart from a number of relatively narrow (in the radial direction) regions in which the applied
perturbation resonates with the equilibrium magnetic field. Magnetic reconnection can take place within the resonant regions to
produce relatively thin magnetic island chains. Within the resonant regions, the plasma response is governed by  linear or nonlinear 
 two-fluid resistive-MHD, depending on whether the induced magnetic island widths are smaller or larger than the corresponding linear layer widths. 
 Thus, when employing the asymptotic matching approach, the equations of marginally-stable ideal-MHD
are solved in the so-called ``outer region'' that comprises most of the plasma (and the surrounding vacuum), the equations of 
linear/nonlinear two-fluid resistive-MHD are solved in the various resonant layers that constitute the so-called ``inner region'', and the two
sets of solutions are then asymptotically matched to one another.

In the nonlinear resonant response regime, the evolution of the island width in a given resonant layer is governed by the familiar Rutherford island width evolution equation.\cite{ruth,ruth1}
On the other hand, the phase evolution of the island chain is conventionally governed by the well-known no-slip constraint.\cite{rfa} According to this constraint,  plasma is trapped inside the
magnetic separatrix of the magnetic island chain,
which forces the chain to co-rotate with the local plasma flow. In reality, however, the plasma
is able to diffuse resistively across the separatrix to some extent.\cite{slip1,slip2}

The primary aim of this paper is to investigate 
what happens when the no-slip constraint is relaxed in a recently-developed model of  RMP-induced
suppression of ELMs in H-mode tokamak discharges.\cite{rftor} The secondary aim is to examine what happens when an improvement is made to the manner in which the resonant
plasma response component of the model interpolates between the linear and the nonlinear regime.

This paper is organized as follows. In Sect.~\ref{s2}, we develop a cylindrical version of
our new resonant plasma response model. In Sect.~\ref{s3},  we use this cylindrical version as a guide for producing a toroidal
version of the model.

\section{Cylindrical Resonant Response Model}\label{s2}
\subsection{Plasma Equilibrium}
Consider a large aspect-ratio, low-$\beta$, tokamak plasma whose magnetic flux surfaces map out (almost)
concentric circles in the poloidal plane.
Such a plasma is well approximated as a periodic cylinder.  Suppose that the minor radius of the plasma
is $a$. Standard cylindrical coordinates ($r$, $\theta$, $z$) are adopted. The system is assumed to
be periodic in the $z$-direction, with periodicity length $2\pi\,R_0$, where $R_0\gg a$ is the simulated
plasma major radius. It is convenient to define the simulated toroidal angle $\phi=z/R_0$.

@@

Equation~(\ref{e5}) is usually coupled with the so-called ``no-slip constraint'':\,\cite{rfa}
\begin{equation}\label{e6}
\frac{d\varphi_s}{dt} = \omega_s.
\end{equation}
The reasoning behind the imposition of this constraint is that the plasma is trapped inside the
magnetic separatrix of the magnetic island chain that forms in the vicinity of the rational surface,
which forces the chain to co-rotate with the local plasma flow. However, in reality, the plasma
is able to diffuse resistively across the separatrix to some extent. Hence, by analogy with Eqs.~(\ref{e3}), (\ref{e4}), and (\ref{e5}), and 
in accordance with Refs.~\onlinecite{slip1} and \onlinecite{slip2}, in this paper we shall modify Eq.~(\ref{e6}) such that
\begin{equation}\label{e7}
\frac{d\varphi_s}{dt} = \omega_s - \frac{E_{sv}}{\tau_R}\,\frac{r_s}{({\cal I}/2)\,W_s}\,\frac{|{\mit\Psi}_v|}{|{\mit\Psi}_s|}\,\sin(\varphi_s-\varphi_v).
\end{equation}
The final term on the right-hand side of the previous equation is the ``nonlinear slip-frequency''. In general,
we would expect this frequency to be relatively small (because it is generally the case that $|\omega_s|\,\tau_R\,(W_s/r_s)\gg 1$
for fully-developed magnetic islands in high-temperature tokamak plasmas\,\cite{rf1,slip2}). Nevertheless, the
nonlinear slip-frequency may be non-negligible for developing island chains, which is why we are including it in our analysis.
(Note that the nonlinear slip-frequency does not appear in rigorous island calculations, such as
that described in Ref.~\onlinecite{rfisland}, and references therein, because such calculations depend crucially on
complicated hierarchical ordering assumptions, according to which the nonlinear slip-frequency is negligible.)

\subsection{Composite Model}
We can combine Eqs.~(\ref{e3}), (\ref{e4}), (\ref{e5}), and (\ref{e7}) to give the following
composite resonant response model that interpolates between the linear and the nonlinear
regimes:
\begin{align}\label{e9}
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\,\frac{d|{\mit\Psi}_s|}{dt} &= E_{ss}\,|{\mit\Psi}_s| + E_{sv}\,|{\mit\Psi}_v|\,\cos(\varphi_s-\varphi_v),\\[0.5ex]
\frac{d\varphi_s}{dt}& = \omega_s - \frac{E_{sv}}{\tau_R}\left(\frac{r_s}{({\cal I}/2)\,W_s+\delta_s}\right)\frac{|{\mit\Psi}_v|}{|{\mit\Psi}_s|}\,\sin(\varphi_s-\varphi_v).\label{e10}
\end{align}
If we define $X_s = |{\mit\Psi}_s|\,\cos\varphi_s$, $Y_s = |{\mit\Psi}_s|\,\sin\varphi_s$, $X_v = |{\mit\Psi}_v|\,\cos\varphi_v$, and $Y_v = |{\mit\Psi}_v|\,\sin\varphi_v$ 
then the previous two equations are more conveniently written in the nonsingular (when $|{\mit\Psi}_s|=0$) forms:\,\cite{slip1,slip2,slip3}
\begin{align}\label{e11}
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\left(\frac{dX_s}{dt}+\omega_s\,Y_s\right)&= E_{ss}\,X_s + E_{sv}\,X_v,\\[0.5ex]
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\left(\frac{dY_s}{dt}-\omega_s\,X_s\right)& = E_{ss}\,Y_s + E_{sv}\,Y_v,\label{e12}
\end{align}
where
\begin{equation}\label{e23x}
W_s = 4\left(\frac{\sqrt{X_s^{\,2}+Y_s^{\,2}}}{s_s\,r_s\,B_\theta(r_s)}\right)^{1/2} r_s.
\end{equation}
With the benefit of hindsight, Eqs.~(\ref{e11}) and (\ref{e12}) could have been derived more directly from the
following heuristic generalization of Eq.~(\ref{e1}):\,\cite{slip2}
\begin{equation}
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\left(\frac{d}{dt}+{\rm i}\,\omega_s\right){\mit\Psi}_s= E_{ss}\,{\mit\Psi}_s+E_{sv}\,{\mit\Psi}_v.
\end{equation}

\section{Modified Toroidal Resonant Response Model}\label{s3}
Our toroidal model of the response of a tokamak plasma to a static, externally generated, RMP is described in
detail in Ref.~\onlinecite{rftor}. In the light of the composite cylindrical resonant response model developed in Sect.~\ref{s2}, we shall
modify the resonant  response component of the aforementioned toroidal  model; this component is described in Appendix~D of Ref.~\onlinecite{rftor}.

Equations~(D1) and (D2) in Ref.~\onlinecite{rftor} are modified such that 
\begin{align}\label{e15}
\left(\hat{W}_k + \hat{\delta}_k\right){\cal S}_k\,\frac{d\hat{\mit\Psi}_k}{d\hat{t}} &= \sum_{k'=1,K}\hat{E}_{kk'}\hat{\mit\Psi}_{k'}\,\cos(\varphi_k-\varphi_{k'}-\xi_{kk'})
+\hat{E}_{kk}\,\hat{\chi}_k\,\cos(\varphi_k-\zeta_k),\\[0.5ex]
{\cal S}_k&=\frac{\tau_R(\hat{r}_k)}{\tau_A},
\end{align}
respectively. Here, 
\begin{align}
\hat{W}_k &=\frac{2\,{\cal I}}{\hat{a}\,\hat{r}_k}\left(\frac{q}{g\,s}\right)_{\hat{r}_k}^{1/2} |\hat{\mit\Psi}_k|^{\,1/2},\\[0.5ex]
\hat{\delta}_k &= \frac{\delta_{\rm SC}(\hat{r}_k)}{R_0\,\hat{a}\,\hat{r}_k},
\end{align}
where $\delta_{\rm SC}(\hat{r})$ is the semi-collisional linear layer width specified in Ref.~\onlinecite{rf1}.

@@

Equation~(D6) of Ref.~\onlinecite{rftor} is modified such that
\begin{align}\label{e19}
\left(\hat{W}_k +\hat{\delta}_k\right){\cal S}_k\left(\frac{d\varphi_k}{d\hat{t}}-\hat{\varpi}_k\right)&=
- \sum_{k'=1,K}\hat{E}_{kk'}\,\hat{\mit\Psi}_{k'}\,\sin(\varphi_k-\varphi_{k'}-\xi_{kk'})-\hat{E}_{kk}\,\hat{\chi}_k\,\sin(\varphi_k-\zeta_k),
\end{align}
where
\begin{equation}
\hat{\varpi}_k= \hat{\varpi}_{k0}+m_k\,{\mit\Delta}\hat{\mit\Omega}_\theta(\hat{r}_k,\hat{t})-n\,{\mit\Delta}\hat{\mit\Omega}_\phi(\hat{r}_k,\hat{t}),
\end{equation}
and $\hat{\varpi}_{k0}= \varpi_{k0}\,\tau_A$, $\hat{\mit\Omega}_\theta= {\mit\Delta}{\mit\Omega}_\theta\,\tau_A$, and $\hat{\mit\Omega}_\phi= {\mit\Delta}{\mit\Omega}_\phi\,\tau_A$. Here, $\varpi_{k0}$
is the so-called ``natural frequency'' at the $k$th rational surface; this quantity is defined as the helical phase velocity of a naturally unstable island chain, resonant at the surface, in the
absence of an RMP (or any other island chains). On the other hand,  ${\mit\Delta}{\mit\Omega}_\theta$ and ${\mit\Delta}{\mit\Omega}_\phi$ are the changes in the plasma
poloidal and toroidal velocity profiles, respectively, induced by the electromagnetic torques that develop in the vicinities of the various rational surfaces in the plasma.
Equation~(\ref{e19}) represents
a major improvement to Eq.~(D6) that takes into  account the finite slip-frequencies of the magnetic island chains induced at
the various rational surfaces in the plasma. In the absence of a finite slip-frequency, the right-hand side of Eq.~(\ref{e19}) would be zero.

If we define $X_k = \hat{\mit\Psi}_k\,\cos\varphi_k$ and $Y_k=\hat{\mit\Psi}_k\,\sin\varphi_k$ then Eqs.~(\ref{e15}) and (\ref{e19}) are
more conveniently written in the nonsingular (when $\hat{\mit\Psi}_k=0$) forms
\begin{align}\label{e21}
\left(\hat{W}_k +\hat{\delta}_k\right){\cal S}_k\left(\frac{dX_k}{d\hat{t}} 
+\hat{\varpi}_k\,Y_k\right)= \sum_{k'=1,K}\hat{E}_{kk'}\,(\cos\xi_{kk'}\,X_{k'}  -\sin\xi_{kk'}\,Y_{k'}) + \hat{E}_{kk}\,\hat{\chi}_k\,\cos\zeta_k,\\[0.5ex]
\left(\hat{W}_k +\hat{\delta}_k\right){\cal S}_k\left(\frac{dY_k}{d\hat{t}} -\hat{\varpi}_k\,X_k\right)
= \sum_{k'=1,K}\hat{E}_{kk'}\,(\cos\xi_{kk'}\,Y_{k'}
+\sin\xi_{kk'}\,X_{k'}) + \hat{E}_{kk}\,\hat{\chi}_k\,\sin\zeta_k,\label{e22}
\end{align}
where
\begin{equation}\label{e23}
\hat{W}_k =\frac{2\,{\cal I}}{\hat{a}\,\hat{r}_k}\left(\frac{q}{g\,s}\right)^{1/2}_{\hat{r}_k} (X_k^{\,2} + Y_k^{\,2})^{\,1/4}.
\end{equation}
Equations (\ref{e15}), (\ref{e19}), (\ref{e21}), (\ref{e22}) and (\ref{e23}) are the toroidal generalizations of Eqs.~(\ref{e9}), (\ref{e10}), (\ref{e11}), (\ref{e12}), and (\ref{e23x}), respectively.

\section{Discussion and Conclusions}

>>>>

\title {Further Modeling of $q_{95}$ Windows for the Suppression of Edge Localized Modes by Resonant Magnetic Perturbations in the DIII-D Tokamak}

@@

\begin{abstract}
An improved resonant plasma response model that describes the interaction between a tokamak plasma and a resonant magnetic perturbation  (RMP) is developed. 
The model interpolates between the linear and the nonlinear response regimes, and takes into account the fact that the slip-frequency
is non-zero in the nonlinear regime. The improved model is incorporated into the {\tt EPEC} toroidal asymptotic matching code.
The modified {\tt EPEC} code is used to model RMP-induced edge-localized-modr mitigation/suppression in DIII-D H-mode discharge \#145380. 
Allowing for a finite slip-frequency (i.e., relaxing the so-called no-slip constraint) is found to slightly facilitate the locking of
driven magnetic island chains to the RMP. The previously obtained conclusion 
that the response of a typical H-mode tokamak plasma to an RMP cannot be accurately modeled by linear theory is confirmed. The previously obtained conclusion that the best agreement between theory and observations is achieved by assuming that the
natural frequencies of tearing modes, in the absence of the RMP, are determined by the local equilibrium ${\bf E}\times{\bf B}$ frequency is also confirmed.
\end{abstract}

\maketitle

\section{Introduction}
In the context of tokamak plasmas, a resonant magnetic perturbation (RMP) refers to an externally
generated, (usually) static, helical magnetic perturbation that resonates at
one or more ``rational'' magnetic flux surfaces within the plasma. Resonant magnetic perturbations have been successfully
employed to compensate error fields (i.e., accidentally produced RMPs), and, thereby,  to prevent the associated formation of
locked modes, in a wide variety of different types of tokamak discharge. \cite{error1,error2,error3} RMPs have also been  successfully used to either  
mitigate or completely suppress  edge-localized-modes (ELMs) in H-mode tokamak discharges.\cite{elm1,elm2,elm3,elm4,elm5,elm6}

``Asymptotic matching'' is by far the most efficient method for modeling the response of a tokamak plasma to an RMP.\cite{fkr,rfa, am1,am2,am3} 
According to the asymptotic matching approach, the response of the plasma to the applied RMP is governed by a combination of flux-freezing and
perturbed force balance---this combination is usually referred to as ``marginally-stable ideal-magnetohydrodynamics (MHD)"---everywhere in the plasma apart from a number of relatively narrow (in the radial direction) regions in which the applied
perturbation resonates with the equilibrium magnetic field. Magnetic reconnection can take place within the resonant regions to
produce relatively narrow magnetic island chains. Within the resonant regions, the plasma response is governed by  linear or nonlinear 
 two-fluid resistive-MHD, depending on whether the induced magnetic island widths are smaller or larger, respectively, than the corresponding linear layer widths. 
 Thus, when employing the asymptotic matching approach, the equations of marginally-stable ideal-MHD
are solved in the so-called ``outer region'' that comprises most of the plasma (and the surrounding vacuum), the equations of 
linear/nonlinear two-fluid resistive-MHD are solved in the various resonant layers that constitute the so-called ``inner region'', and the two
sets of solutions are then asymptotically matched to one another.

In the nonlinear resonant response regime, the evolution of the island chain width in a given resonant layer is governed by the familiar ``Rutherford island width evolution equation''.\cite{ruth,ruth1}
On the other hand, the phase evolution of the island chain is conventionally governed by the well-known ``no-slip constraint''.\cite{rfa} According to this constraint,  plasma is trapped inside the
magnetic separatrix of the island chain,
which forces the chain to co-rotate with the local plasma flow. In reality, however, the plasma
is able to diffuse resistively across the separatrix to some extent.\cite{slip1,slip2}

The recently developed, toroidal, asymptotic matching code {\tt EPEC} has been used to simulate 
RMP-induced ELM mitigation/suppression in DIII-D H-mode plasmas.\cite{rftor,rftor1}
The original {\tt EPEC} model incorporates the no-slip constraint.
The first aim of this paper is to investigate 
what happens when the no-slip constraint is relaxed in the aforementioned {\tt EPEC} simulations.
 The second aim is to examine what happens when an improvement is made to the manner in which the resonant
plasma response component of the {\tt EPEC} model interpolates between the linear and the nonlinear regimes.

\section{Cylindrical Resonant Response Model}\label{s2}
\subsection{Introduction}
For the sake of simplicity, and also to make a better connection with previous research, let us commence our investigation by discussing
the resonant response of a tokamak plasma to an RMP in cylindrical geometry.

\subsection{Plasma Equilibrium}
Consider a large aspect-ratio, low-$\beta$, tokamak plasma whose magnetic flux surfaces map out (almost)
concentric circles in the poloidal plane.
Such a plasma is well approximated as a periodic cylinder.  Suppose that the minor radius of the plasma
is $a$. Standard cylindrical coordinates ($r$, $\theta$, $z$) are adopted. The system is assumed to
be periodic in the $z$-direction, with periodicity length $2\pi\,R_0$, where $R_0\gg a$ is the simulated
plasma major radius. It is convenient to define the simulated toroidal angle $\phi=z/R_0$.

@@

Equation~(\ref{e5}) is usually coupled with the so-called ``no-slip constraint'':\,\cite{rfa}
\begin{equation}\label{e6}
\frac{d\varphi_s}{dt} = \omega_s.
\end{equation}
The reasoning behind the imposition of this constraint is that the plasma is trapped inside the
magnetic separatrix of the  island chain that forms in the vicinity of the rational surface,
which forces the chain to co-rotate with the local plasma flow. However, in reality, the plasma
is able to diffuse resistively across the separatrix to some extent. Hence, by analogy with Eqs.~(\ref{e3}), (\ref{e4}), and (\ref{e5}), and 
in accordance with Refs.~\onlinecite{slip1} and \onlinecite{slip2}, in this paper we shall modify Eq.~(\ref{e6}) such that
\begin{equation}\label{e7}
\frac{d\varphi_s}{dt} = \omega_s - \frac{E_{sv}}{\tau_R}\,\frac{r_s}{({\cal I}/2)\,W_s}\,\frac{|{\mit\Psi}_v|}{|{\mit\Psi}_s|}\,\sin(\varphi_s-\varphi_v).
\end{equation}
The final term on the right-hand side of the previous equation is the ``nonlinear slip-frequency''. In general,
we would expect this frequency to be relatively small (because it is generally the case that $|\omega_s|\,\tau_R\,(W_s/r_s)\gg 1$
for fully-developed magnetic islands in high-temperature tokamak plasmas\,\cite{rf1,slip2}). Nevertheless, the
nonlinear slip-frequency may be non-negligible for developing island chains, which is why we are including it in our analysis.
(Note that the nonlinear slip-frequency does not appear in rigorous island calculations, such as
that described in Ref.~\onlinecite{rfisland}, and references therein, because such calculations depend crucially on
complicated hierarchical ordering assumptions, according to which the nonlinear slip-frequency is negligible.)

\subsection{Composite Model}
We can combine Eqs.~(\ref{e3}), (\ref{e4}), (\ref{e5}), and (\ref{e7}) to give the following
composite resonant response model that interpolates between the linear and the nonlinear
regimes:
\begin{align}\label{e9}
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\,\frac{d|{\mit\Psi}_s|}{dt} &= E_{ss}\,|{\mit\Psi}_s| + E_{sv}\,|{\mit\Psi}_v|\,\cos(\varphi_s-\varphi_v),\\[0.5ex]
\frac{d\varphi_s}{dt}& = \omega_s - \frac{E_{sv}}{\tau_R}\left(\frac{r_s}{({\cal I}/2)\,W_s+\delta_s}\right)\frac{|{\mit\Psi}_v|}{|{\mit\Psi}_s|}\,\sin(\varphi_s-\varphi_v).\label{e10}
\end{align}
If we define $X_s = |{\mit\Psi}_s|\,\cos\varphi_s$, $Y_s = |{\mit\Psi}_s|\,\sin\varphi_s$, $X_v = |{\mit\Psi}_v|\,\cos\varphi_v$, and $Y_v = |{\mit\Psi}_v|\,\sin\varphi_v$ 
then the previous two equations are more conveniently written in the non-singular (when $|{\mit\Psi}_s|=0$) forms:\,\cite{slip1,slip2,slip3}
\begin{align}\label{e11}
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\left(\frac{dX_s}{dt}+\omega_s\,Y_s\right)&= E_{ss}\,X_s + E_{sv}\,X_v,\\[0.5ex]
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\left(\frac{dY_s}{dt}-\omega_s\,X_s\right)& = E_{ss}\,Y_s + E_{sv}\,Y_v,\label{e12}
\end{align}
where
\begin{equation}\label{e23x}
W_s = 4\left(\frac{\sqrt{X_s^{\,2}+Y_s^{\,2}}}{s_s\,r_s\,B_\theta(r_s)}\right)^{1/2} r_s.
\end{equation}
With the benefit of hindsight, Eqs.~(\ref{e11}) and (\ref{e12}) could have been derived more directly from the
following heuristic generalization of Eq.~(\ref{e1}):\,\cite{slip2}
\begin{equation}
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\left(\frac{d}{dt}+{\rm i}\,\omega_s\right){\mit\Psi}_s= E_{ss}\,{\mit\Psi}_s+E_{sv}\,{\mit\Psi}_v.
\end{equation}

\section{Modified Toroidal Resonant Response Model}\label{s3}
Let us now generalize our investigation to toroidal geometry.

The toroidal model of the response of a tokamak plasma to a static, externally generated, RMP that is implemented in the {\tt EPEC} code\,\cite{rftor1} is described in
detail in Ref.~\onlinecite{rftor}. In the light of the composite cylindrical resonant response model developed in Sect.~\ref{s2}, we propose to 
modify the resonant  response component of the {\tt EPEC} model (this component is described in Appendix~D of Ref.~\onlinecite{rftor}).

Equations~(D1) and (D2) in Ref.~\onlinecite{rftor} are modified such that 
\begin{align}\label{e15}
\left(\hat{W}_k + \hat{\delta}_k\right){\cal S}_k\,\frac{d\hat{\mit\Psi}_k}{d\hat{t}} &= \sum_{k'=1,K}\hat{E}_{kk'}\,\hat{\mit\Psi}_{k'}\,\cos(\varphi_k-\varphi_{k'}-\xi_{kk'})
+\hat{E}_{kk}\,\hat{\chi}_k\,\cos(\varphi_k-\zeta_k),\\[0.5ex]
{\cal S}_k&=\frac{\tau_R(\hat{r}_k)}{\tau_A},
\end{align}
respectively. Here, 
\begin{align}
\hat{W}_k &=\frac{2\,{\cal I}}{\hat{a}\,\hat{r}_k}\left(\frac{q}{g\,s}\right)_{\hat{r}_k}^{1/2} |\hat{\mit\Psi}_k|^{\,1/2},\\[0.5ex]
\hat{\delta}_k &= \frac{\delta_{\rm SC}(\hat{r}_k)}{R_0\,\hat{a}\,\hat{r}_k},
\end{align}
where $\delta_{\rm SC}(\hat{r})$ is the semi-collisional linear layer width specified in Ref.~\onlinecite{rf1}.

@@

Equation~(D6) of Ref.~\onlinecite{rftor} is modified such that
\begin{align}\label{e19}
\left(\hat{W}_k +\hat{\delta}_k\right){\cal S}_k\left(\frac{d\varphi_k}{d\hat{t}}-\hat{\varpi}_k\right)&=
- \sum_{k'=1,K}\hat{E}_{kk'}\,\hat{\mit\Psi}_{k'}\,\sin(\varphi_k-\varphi_{k'}-\xi_{kk'})-\hat{E}_{kk}\,\hat{\chi}_k\,\sin(\varphi_k-\zeta_k),
\end{align}
where
\begin{equation}
\hat{\varpi}_k= \hat{\varpi}_{k\,0}+m_k\,{\mit\Delta}\hat{\mit\Omega}_\theta(\hat{r}_k,\hat{t})-n\,{\mit\Delta}\hat{\mit\Omega}_\phi(\hat{r}_k,\hat{t}),
\end{equation}
and $\hat{\varpi}_{k\,0}= \varpi_{k\,0}\,\tau_A$, $\hat{\mit\Omega}_\theta= {\mit\Delta}{\mit\Omega}_\theta\,\tau_A$, and $\hat{\mit\Omega}_\phi= {\mit\Delta}{\mit\Omega}_\phi\,\tau_A$. Here, $\varpi_{k\,0}$
is the so-called ``natural frequency'' at the $k$th rational surface; this quantity is defined as the helical phase velocity of a naturally unstable island chain, resonant at the surface, in the
absence of an RMP (or any other island chains). On the other hand,  ${\mit\Delta}{\mit\Omega}_\theta$ and ${\mit\Delta}{\mit\Omega}_\phi$ are the changes in the plasma
poloidal and toroidal velocity profiles, respectively, induced by the electromagnetic torques that develop in the vicinities of the various rational surfaces in the plasma.
Equation~(\ref{e19}) represents
a major improvement to Eq.~(D6) that takes into  account the finite slip-frequencies of the magnetic island chains induced at
the various rational surfaces in the plasma. In the absence of a finite slip-frequency, the right-hand side of Eq.~(\ref{e19}) would be zero.

If we define $X_k = \hat{\mit\Psi}_k\,\cos\varphi_k$ and $Y_k=\hat{\mit\Psi}_k\,\sin\varphi_k$ then Eqs.~(\ref{e15}) and (\ref{e19}) are
more conveniently written in the non-singular (when $\hat{\mit\Psi}_k=0$) forms
\begin{align}\label{e21}
\left(\hat{W}_k +\hat{\delta}_k\right){\cal S}_k\left(\frac{dX_k}{d\hat{t}} 
+\hat{\varpi}_k\,Y_k\right)= \sum_{k'=1,K}\hat{E}_{kk'}\,(\cos\xi_{kk'}\,X_{k'}  -\sin\xi_{kk'}\,Y_{k'}) + \hat{E}_{kk}\,\hat{\chi}_k\,\cos\zeta_k,\\[0.5ex]
\left(\hat{W}_k +\hat{\delta}_k\right){\cal S}_k\left(\frac{dY_k}{d\hat{t}} -\hat{\varpi}_k\,X_k\right)
= \sum_{k'=1,K}\hat{E}_{kk'}\,(\cos\xi_{kk'}\,Y_{k'}
+\sin\xi_{kk'}\,X_{k'}) + \hat{E}_{kk}\,\hat{\chi}_k\,\sin\zeta_k,\label{e22}
\end{align}
where
\begin{equation}\label{e23}
\hat{W}_k =\frac{2\,{\cal I}}{\hat{a}\,\hat{r}_k}\left(\frac{q}{g\,s}\right)^{1/2}_{\hat{r}_k} (X_k^{\,2} + Y_k^{\,2})^{\,1/4}.
\end{equation}
Equations (\ref{e15}), (\ref{e19}), (\ref{e21}), (\ref{e22}) and (\ref{e23}) are the toroidal generalizations of Eqs.~(\ref{e9}), (\ref{e10}), (\ref{e11}), (\ref{e12}), and (\ref{e23x}), respectively.

\section{EPEC Modeling of DIII-D Discharge \#145380}\label{s4}
\subsection{Introduction}
DIII-D discharge \#145380 is an ITER-Similar-Shape (ISS), ELMing, H-mode discharge, with a toroidal magnetic field $B_T=-1.9\,{\rm T}$, in which the
plasma current, $I_p$, is slowly ramped over a 2 second interval in order to scan the magnetic safety-factor.\cite{d3d,d3d2} The majority ions are Deuterium, whereas the
minority ions are Carbon~VI.
Figure~\ref{fig1} gives an overview of this discharge. ${\mit\Psi}_N$ is the conventional normalized poloidal magnetic flux.
A static RMP is applied to the plasma by running steady $n=3$ currents through the I-coil system.\cite{icoil}
Three windows of ELM suppression or mitigation are evident: the first (in which ELMs are mitigated, but not
entirely suppressed) extends from $t=2840$--$2980$ ms; the
second (in which ELM are entirely suppressed) extends from $t=3320$--$3560$ ms; and the third (in which ELMs are
entirely suppressed) extends from $t=3880$--$4200$ ms. The 
first, second, and third ELM-suppression/mitigation windows extend over the $\overline{q_{95}}$ ranges $3.88$--$3.94$, $3.62$--$3.73$,
and $3.33$--$3.48$, respectively. Here, $\overline{q_{95}}$ is least-squares linear fit to the variation of the safety-factor at the ${\mit\Psi}_N = 0.95$ flux surface with time.

\subsection{Relaxation of No-Slip Constraint}\label{s4a}
As is discussed in Ref.~\onlinecite{rftor1}, the best agreement between the predictions of the {\tt EPEC} code and experimental observations is
obtained by assuming that the natural frequencies at the various rational surfaces in the plasma are determined by the
local equilibrium ${\bf E}\times {\bf B}$ frequencies. In other words, 
\begin{equation}\label{e24}
\varpi_{k\,0} =-n\,\omega_E({\mit\Psi}_{p\,k}),
\end{equation}
where 
\begin{equation}
\omega_E({\mit\Psi}_p) = -\frac{d{\mit\Phi}}{d{\mit\Psi}_p}.
\end{equation}
Here, ${\mit\Phi}({\mit\Psi}_p)$ is the equilibrium scalar electric potential, ${\mit\Psi}_p$ the equilibrium poloidal magnetic flux (divided by $2\pi$), 
and ${\mit\Psi}_{p\,k}$  the value of ${\mit\Psi}_p$ at the $k$th rational surface.  Equation~(\ref{e24}) holds at
every rational surface in all of the simulations discussed in Sect.~\ref{s4a}.

\subsubsection{No-Slip Constraint Imposed}
In our first simulation, the no-slip constraint is imposed at all rational surfaces in the plasma. This implies that the resonant response
model at the $k$th rational surface consists of Eq.~(\ref{e15}) and Eq.~(\ref{e19}), with the right-hand side of the latter equation
set to zero; in other words,  the slip-frequency is set to zero at the $k$th rational surface.

The upper panel of Fig.~\ref{fig2} shows the natural frequencies, in the absence of the RMP, of the various  $n=3$ tearing modes 
 that are resonant in the pedestal of DIII-D discharge \#145380. It can be seen that 
the natural frequencies of the $m=11$ (brown), the $m=10$ (magenta), and the $m=9$ (cyan), tearing modes have their
zero-crossings in the  first, second, and third, ELM-suppression/mitigation window, respectively. 
  The lower panel of Fig.~\ref{fig2} shows the natural frequencies, in the presence of the RMP, of the various $n=3$
tearing modes that are resonant in the pedestal. It is apparent that if the magnitude of the natural frequency in the absence of the RMP falls below about
$3\,{\rm krad/s}$ then the associated tearing mode locks to the RMP (i.e., its true natural frequency becomes zero).

The upper panel of Fig.~\ref{fig3} shows the widths of the various vacuum magnetic island chains driven by the applied $n=3$ RMP in the pedestal region as functions of $
\overline{q_{95}}$.
The vacuum island widths are the  expected island widths in the absence of any shielding of driven magnetic reconnection due to
plasma flow; in other words, they are the  widths of the island chains that would be driven if all of the natural frequencies were zero. It can be seen that
the driven magnetic island chains extend over most of the pedestal, implying a very significant degradation in pedestal energy and particle confinement due to the
expected flattening of the density and temperature profiles across the chains.\cite{flat} Moreover, the driven island chains at the
bottom of the pedestal overlap, implying that the magnetic field in this region is rendered stochastic. The lower panel
of Fig.~\ref{fig3} shows the actual widths of the various magnetic island chains driven by the applied $n=3$ RMP as functions of time. It is apparent 
that the true island widths are generally much smaller than the vacuum island widths, as a consequence of the shielding of driven magnetic
reconnection due to plasma flow. However, the shielding breaks down at various rational surfaces when the associated natural frequency is zero (see the lower panel of
Fig.~\ref{fig3}). In this situation, the driven island width is similar to the vacuum island width.

According to the prevailing hypothesis, RMP-induced ELM mitigation/suppression in an H-mode tokamak discharge occurs when a comparatively wide, locked, magnetic
island chain is generated at the top of the pedestal.\cite{hu} The temperature and density flattening associated with the chain
is presumed to limit the radial expansion of the
pedestal, and, thereby, prevent it from attaining a width sufficient to destabilize peeling-ballooning modes\,\cite{connor}  (which are thought to trigger ELMs).\cite{d3d}
It has also been suggested that  the well-known ``density pump-out" phenomenon, which is often observed when RMPs are applied
to H-mode tokamak plasmas, is associated with the formation of overlapping island chains
at the bottom of the pedestal.\cite{hu}

According to Fig.~\ref{fig3},  
 the comparatively wide, $m=11$ (brown),  $m=10$ (magenta), and  $m=9$ (cyan), locked magnetic island chains which are predicted to be driven by the RMP in the pedestal of DIII-D discharge \#145380 line up
fairly well with the first, second, and third, ELM-suppression/mitigation windows, respectively. To be more exact, the island chains are located at
the top of the pedestal (i.e., ${\mit\Psi}_N=0.925$), and also occur in roughly the correct time intervals. This is consistent with the hypothesis that the formation of locked
magnetic island chains at the top of the pedestal causes ELM mitigation/suppression. 
Note that overlapping island chains
are also driven at the bottom of the pedestal; as has already been mentioned, such island chains are associated with the density pump-out phenomenon.
